API design for machine learning: experiences from scikit-learn

### 2014-02-18 paper.tex
"generalize" introduction

The \sklearn library provides a comprehensive suite of machine learning tools
for Python.
It extends this general-purpose programming language with machine learning
operations: learning algorithms, preprocessing tools, model selection procedures
and a composition mechanism to produce complex ML workflows.
The ambition of the library is to provide efficient implementations
of well-established algorithms within a
programming environment that is accessible to non-experts and
reusable in various scientific areas.
The library is distributed under the simplified BSD license
to encourage its use in both academia and industry.

\sklearn is designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries
\citep{oliphant2007python, vanderwalt2011}.
NumPy augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy extends it further with common numerical operations,
either by implementing these in Python/NumPy
or by wrapping existing C/C{}\verb!++!/Fortran code.

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api}
the central concepts of our application programming interface (API).
In section~\ref{sec:advanced-api}, we discuss
advanced API mechanisms built on the core interface,
focusing on composition of operations
and on the important problem of model selection.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the objectives for
a \sklearn 1.0 release.
We conclude by summarizing the major points of this paper in
section~\ref{sec:conclusions}.

>>>>

In this paper, we present a high-level application programming interface (API)
for machine learning tasks.
The API presented here is primarily that of \sklearn \citep{pedregosa2011},
an open~source Python library
that provides a comprehensive suite of machine learning tools,
extending the ``ecosystem'' of scientific Python tools
centered around NumPy and SciPy \citep{oliphant2007python, vanderwalt2011}.
\sklearn's API principles have been adopted by other packages including
\textit{astroML} \citep{vanderplas2012astroML},
a machine learning toolkit for astronomy;
\textit{wiseRF}\footnote{\url{http://wise.io}},
a commercial random~forest implementation;
and \textit{lightning},\footnote{\url{https://github.com/mblondel/lightning}}
the source code of a recently-proposed sparse multiclass
algorithm \citep{mblondel-mlj2013}.
% XXX PyStruct, seqlearn?
While code examples in this article are in Python and use \sklearn,
the principles exposed are general
and should apply in any modern programming language.

In this paper, we present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api}
the central concepts of our application programming interface (API).
In section~\ref{sec:advanced-api}, we discuss
advanced API mechanisms built on the core interface,
focusing on composition of operations
and on the important problem of model selection.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the objectives for
a \sklearn 1.0 release.
We conclude by summarizing the major points of this paper in
section~\ref{sec:conclusions}.
